---
title: "How to speed up R code"
author: "Fibears"
date: "2015年10月17日"
output:
  beamer_presentation:
    colortheme: dolphin
    fig_caption: yes
    fonttheme: structurebold
    highlight: default
    includes:
      in_header: header.tex
    pandoc_args: --latex-engine=xelatex
    slide_level: 2
    theme: AnnArbor
  ioslides_presentation:
    highlight: default
---
## Content
1. Abstract
2. Your Computer
3. One R instance, one core
4. Using the other cores of your computer

## Abstract
* **R** is more and more popular in some fields such as Statistics. But the users then 
faces total execution times of his codes that are hard to work with: hours,days,
even weeks.
* **Goal**: How to reduce the total execution time of your R codes

## Chap 1.
* Your Computer
    + 1.1 HDD (The hard disk drive)
    + 1.2 RAM (The random-access memory)
    + 1.3 CPU (The central processing unit)

## 1.1 HDD
* The hard disk drive is where all the data of the computer is stored,even 
when it is powered down.
* Cheap to produce
* But,they are slow!

```{r eval=FALSE}
system.time({
    data <- rnorm(10^7)
    write.table(data, "data.txt")
})
```

## 1.1 HDD
* This file _data.txt_ was 280.5 MBs,so we can roughly infer that a 
4-GB dataset will require 500 seconds to be written down on a HDD.
* A rule when writing R code: only access the HDD at the beginning of 
your code(to load some data) and at the end(to save some results).
* Use SSD: You can store files you have to access quite often on the 
small SSD and files of lesser interest on the large HDD.

## 1.2 RAM
* The random-access memory is also used to store files.
    + read.table(): The dataset is copied from the HDD(or SSD) and pasted 
    on the RAM.
    + RAM is a much faster data storage technology, but everything stored 
    on the RAM is lost when the computer is powered down.
    + Expensive
* 32-bit R or 64-bit R?
    + 32-bit programs cannot use more than 2 GBs of RAM.
    + 64-bit programs are allowed to use as much as RAM as they need.

~~~~~~~
R version 3.2.2 (2015-08-14) -- "Fire Safety"
Copyright (C) 2015 The R Foundation for Statistical Computing
Platform: x86_64-apple-darwin13.4.0 (64-bit)
~~~~~~~

## 1.3 CPU
* CPU: The central processing unit--the heart of your computer
* Multi-core: __R__ is not natively able to use several cores at the 
same time! But,some packages allowing __R__ to use more than one core 
at the same time. 

## Chap 2.
* One R instance, one core
    + 2.1 Writing your code right
    + 2.2 What part of my code is slow?
    + 2.3 Writing part of your code in C

## 2.1 Writing your code right
Two ways to measure how fast your code currently run.
```{r eval=FALSE}
time.start <- proc.time()[[3]]
#### Your code ####
time.end <- proc.time()[[3]]
time.end - time.start
```
```{r eval=FALSE}
system.time({
    ### Your code ###
})[[3]]
```

## 2.1 Writing your code right
* R works better with vectors, matrics and tables. Loops are, in 
general, more expensive, as well as dataframes and lists.
    + Example 1.
    ```{r}
    n <- 10000
    system.time({
    a <- vector(length=n); for(i in 1:n) a[i]=2*pi*sin(i)
    })
    system.time({
    b <- NULL; for(i in 1:n) b[i]=2*pi*sin(i)
    })
    ```


## 2.2 What part of my code is slow?


## 2.3 Writing part of your code in C




## Chap 3.
* Using the others cores of your computer
    + 3.1 Several R instances, one core for each
    + 3.2 One R instance, several cores

## 3.1 Several R instances, one core for each


## 3.2 One R instance, several cores




