---
title: "Lecture 2 SCR"
subtitle: Main Classes and Functions
runtime: shiny
output: ioslides_presentation
---



## Recap of week 1

- Relevance of R

- R basic structures 

- Reproducible report


# Topics of week 2:| Vectors, Matrices, Arrays and Lists (Ch. 2 - 4)

##Content of today

**11.15 - 12.00: Vectors and Matrices and Arrays (1st half)**   


**12.15 - 13.00: Lab meeting**   

  
**13.45 - 14.30: Matrices and Arrays (2nd half) and Lists**   


**14.45 - 15.30: Lab meeting and Peer-review Assignment**   


## Week 2 learning objectives 


- construct vectors in several ways (using ```c()```, ```:```, ```seq()```, ```rep()```, ```paste()```, ```numeric()```, and ```rnorm()```)
- construct other data structures from vectors
- solve computation problems in R with vectors, matrices, arrays, lists and dataframes
- get insight in the ```*apply()``` family functions
- write R functions with vectors and/or matrices as input arguments and a list as output


# Chapter 2: | Vectors

Creating vectors in R   

Vectorization   


## Introduction Vectors 

- **Creating vectors in R** (par. 2.1, 2.4, swirl lesson 2)   

- Vector indexing (self-study par. 2.4.2)   

- Recycling (par 2.3; swirl lesson 1)    

- **Vectorization** (par 2.6)   

- Missing values (par 2.7; swirl lesson 4)  

- Filtering (par 2.8; swirl lesson 5 "subsetting vectors")   


## Creating vectors in R (```typeof```)

Use ```c()``` to generate numeric vectors (type = "double") or character vectors 

```{r, label = cfunction_ex2}
x1 <- c(1, 3, "a", "b"); 
class(x1)
typeof(x1)
```


## Creating vectors in R (```typeof```)

Use the ```:``` Operator to generate numeric vectors of type integer or double.

```{r, label=operator_ex1}
x2 <- 1.5:10.5
typeof(x2)
```
   
   
More about ```typeof()``` vs ```class()```? [See  here](http://stackoverflow.com/questions/6258004/r-types-and-classes-of-variables)   



## Creating vectors in R (continued)

Use the function ```seq()``` or ```rep()``` (swirl lesson 2);   

Use function ```paste()``` to concatenate vector elements without spaces in between:

```{r}
x3 <- paste(c("a", "b", "c", "d"), collapse = "")
x3
```
or, for example, to create a vector with variable names:

```{r}
x4 <- paste("X", 1:5, sep = "")
x4
```


## Creating vectors in R (continued)

Use ```rnorm()``` to generate a random normally distributed variable, for example, with $\mu=10$ and $\sigma=2$ or $\sigma=1$:   

```{r}
set.seed(350)
x5 <- rnorm(200, mean = 10, sd = 2)
x6 <- rnorm(200, 10, 1)
```


## Creating vectors in R (continued)

Use ```rnorm()``` to generate a random normally distributed variable, for example, with $\mu=10$ and $\sigma=2$ or $\sigma=1$:   


```{r, fig.height = 3.5, fig.width = 8, echo = FALSE}
par(mfrow = c(1, 2))
hist(x5, xlim = c(4, 16), ylim = c(0, 60), breaks = 11, main = "Histogram of x5")
hist(x6, xlim = c(4, 16), ylim = c(0, 60), breaks = 11, main = "Histogram of x6")
```


## Vectorization

Function that is *vectorized*: a function is applied individually (and memoryless) to each element of a vector.

Examples of vectorized functions:

The functions/operators ```+```, ```*```, ```/```, ```>``` , ```==```, ```&```, ```|``` are good examples. (see book 2.4.1, 2.6.1, 2.10) 
```{r}
x <- c(1, 6, 2)
y <- c(3, 7, 1)
x > 2 & y > 2
x[x > 2 & y > 2]
```

## Vectorized functions:

- The functions ```sqrt()```, ```round()```, ```ifelse()``` (see also 2.6.1, 2.9)

```ifelse()``` changes a logical vector into a character or numeric vector:
```{r}
x <- 1:8
y <- ifelse(x%%2 == 1, "odd", "even")
y
z <- ifelse(x%%2 == 1, 1, 0)
z
```

## Question time...

Which functions can you use to create a vector?

What is an easy way to generate a vector with 10 zero's?

Which of the below examples does not work? 

```x1 <- 1.3 : 10.3```

```x2 <- a : z```

```x3 <- seq(1 : 5, length = 9)```


## Question time continued

Given:
```{r, eval = FALSE}
x <- c(6, 3, 9)
y <- c(3, 3, 10)
```

Which of the following code is not a vectorized operation?

```
x + 4

x %*% x

mean(x)

x > 4 & y > 4
```

# Chapter 3: | Matrices and Arrays

Matrix indexing and filtering   

Applying functions to matrix rows and columns   

Adding and deleting matrix rows and columns   

Unintended dimension reduction   



## Matrices

- Creating matrices (self-study par. 3.1, swirl lesson 6)
- Matrix algebra (self-study par. 3.2.1; course linear models)
- **Matrix indexing and filtering** (par. 3.2.2 and 3.2.3)
- **Applying functions to matrix rows and columns** (par. 3.3)
- **Adding and deleting matrix rows and columns** (par. 3.4)
- Vector / matrix distinction (self-study par. 3.5)
- **Unintended dimension reduction** (par. 3.6)


## Matrix indexing

Given a matrix Z:
```{r, label = matZ, echo = FALSE}
Z <- matrix(c(4, 6.5, 10, 8, 0, 1, 1, 0), ncol = 2)
colnames(Z) <- c("test", "group")
Z
```
What is the test score of student 4?
```{r}
Z[4, "test"]
```


## Matrix filtering 

Create a vector "index", with the row numbers of Z:
```{r}
index <- 1:nrow(Z) 
```
Then make a selection of the row numbers you are interested in (e.g. group 0)
```{r}
index0 <- index[Z[, 2] == 0]
index0
```

----

Then make a selection of the row numbers you are interested in (e.g. group 0)
```{r}
index0 <- index[Z[, 2] == 0]
index0
```

Use this index0 in your calculations for group 0, for example:
```{r}
mean(Z[index0, 1])
```



## Sorting a column of a matrix

```{r}
sort(Z[, 1]); 
```
```{r, eval = FALSE}
Z[sort(Z[, 1]), ]# does not work!
```

## Sort a matrix by a column

```{r}
indext <- order(Z[, 1]); 
indext
Z[indext, ]
```


## Matrix row/column functions

*In general:*

To apply a function (```f```) on rows of a matrix (```Z```): Use ```apply(Z, 1, f)```   

To apply a function on the columns of a matrix: Use ```apply(Z, 2, f)```   

*More specific:* 

Use existing functions:

- ```rowSums()``` and ```colSums()```;

- ```rowMeans()``` and ```colMeans()```.


## Examples with apply()
```{r, echo = FALSE}
Z <- matrix(4:9, ncol=2)
Z
```

```{r}
f <- function(x) {
  xnew <- x - mean(x)
}
apply(Z, 2, f)
```

## Note on apply

If the function to be applied returns a vector of *k* components, then the result of apply() will have *k* rows.

Example: Apply function f to the rows of matrix Z:
```{r, echo = FALSE}
Z <- matrix(4 : 9, ncol = 2)
Z
```
```{r}
apply(Z, 1, f)
```

## Strategy to read book

- Read the example (without performing the code);

- Try to understand the R code (imagine what the output is, discuss it with your neighbour);

- Perform the example in R (download the [Code from the book](http://www.nostarch.com/artofr.htm))


Apply this strategy to Example 3.3.2 of the book pp. 72 - 73
```{r, echo = FALSE}
findols <- function(x) {
   findol <- function(xrow) {
       mdn <- median(xrow)
       devs <- abs(xrow-mdn)  
       return(which.max(devs))  
   }
   return(apply(x, 1, findol))  
}
```
Use the following matrix as input:
```{r}
rs <- matrix(c(150, 250, 5, 200, 200, 100, 600, 150, 125, 175, 50, 
    75, 60, 200, 65), nrow = 3, byrow = TRUE)
```


## Solution

```{r}
rs
findols(rs)
```


## Extra: add variable names to a matrix

A vector with variable names and a matrix Z:
```{r}
x7 <- paste("X", 1:5, sep = "")
x7
Z <- matrix(0, nrow = 10, ncol = 5)
colnames(Z) <- x7
Z[1, ]
```


## Adding rows

Use the function rbind() to combine matrices by rows:
```{r}
Y <- matrix(c(1:6), nrow = 2)
V <- matrix(c(7:15), nrow = 3)
rbind(Y, V)
```

The number of columns?/rows? of Y and V should be equal


## Adding columns

Use the function cbind() to combine matrices by columns:
```{r}
Y <- matrix(c(1:6), nrow = 2)
W <- matrix(c(7:16), nrow = 2)
cbind(Y, W)
```
The number of columns?/rows? of Y and W should be equal


## (Unintended) Dimension reduction

If you select a row from a matrix, the result is NOT a matrix: the two dimensions are reduced to one. For example:
```{r}
store2 <- rs[2, ]
dim(store2)
```

To keep the two dimensions use:
```{r}
store2 <- rs[2, , drop = FALSE]  #empty space between the comma's: take the complete row (i.e., all columns) 
dim(store2)
```

# Labmeeting

**Break (15 minutes)**

**Labmeeting week 2**

Exercises of the labmeeting are on Blackboard 



# Part 2 | Arrays, Lists and Data frames


# Chapter 3 (continued):| Arrays 

Higher dimensional arrays   


## Higher-dimensional arrays

A matrix is a special case of an array (i.e., a two-dimensional array). Arrays are multidimensional. For example, create a three-dimensional array with a vector as input:

```{r, label = cre_A3}
A3 <- array(seq(1, 16), dim = c(2, 4, 2)); A3
```


## Array indexing

Order of the dimensions is: row, column, layer.

What are the values for the second row on the second layer of "example"?

```{r, label = index_A3}
A3[2, , 2]
```

Read pages 82 and 83. Discover the typo in the matrix of firsttest on p. 82.


## Array indexing (dim. red.)

```{r}
class(A3[2, , 2])
class(A3[2, , 2, drop = FALSE])
```
```{r, eval = FALSE}
apply(A3, 3, sum)
```

## Array indexing (dim. red.)

```{r, label = index1_A3}
class(A3[2, , 2])
class(A3[2, , 2, drop = FALSE])
apply(A3, 3, sum)
```





# Chapter 4: | Lists

Creating lists and properties of a list   

List indexing, components and values   

Applying functions to lists   



## What is a list?

A list structure can combine objects of different types/modes. Each object is called a "component" of a list.

```{r}
info <- list(teacher = c("Maarten", "Elise"), week = 1:7)
info
```


## A recursive list: a list within a list

A list can be regarded as a vector of components. It can also combine ```list()``` objects (par. 4.5) 

```{r, label = cre_obj2}
SCR <- vector('list', 2)
names(SCR) <- paste0("year", 1:2)
SCR[[1]] <- list(year = "2014/2015", info = info)
```

## What are the components of a list?

How many components?
```{r, eval = FALSE}
length(SCR)
```

What are the names?
```{r, eval = FALSE}
attributes(SCR)
```

Is there more?
```{r, eval = FALSE}
str(SCR)
```


## List components

How many components?
```{r, eval = TRUE}
length(SCR)
```

What are the names?
```{r, eval = TRUE}
attributes(SCR)
```

## List components

Is there more?
```{r, eval = TRUE}
str(SCR)
```

## Get rid of ```list```

```{r}
unlist(SCR)
```




## Example of existing lists in R

A list is often used as output of a function. For example:

```{r}
lm1 <- lm(dist ~ speed, data = cars)
attributes(lm1)
```

## Ask for the specific components of a list

```{r}
lm1$coefficients  #gives the coefficients
lm1[[1]]  #gives the coefficients
lm1[[2]][1:5]  #gives the first five values of residuals
```


## Do try it at home

```{r}
f <- function(x) {
    mx <- mean(x)
    xnew <- x - mean(x)
    obj <- list(mean = mx, dfm = xnew)
    return(obj)
}
obj3 <- f(c(1, 2, 3))
obj3
```

## List indexing

If you use double brackets to ask for a component of a list: the component is returned (having the class of the object). 

```{r}
obj3[[2]]; 
```

If you use single brackets to ask for a component of a list, the result is a list.

```{r}
obj3[2]
```



## Question time... (a bit advanced...)

Can anyone explain why

```{r, label = cre_SCRwrong}
SCR <- vector('list', 2)
SCR[1] <- list(year = "2014/2015", info = info); 
```
gives a warning, as compared to

```{r, label = cre_SCRright}
SCR <- list(); length(SCR) <- 2
SCR[[1]] <- list(year = "2014/2015", info = info); 
```

?


## Applying functions to lists: lapply()

Use lapply() to perform a function on components of a list, for example, z was given by:
```{r, echo = FALSE}
obj3
```
```{r, label = ex_lapply}
lapply(obj3, length) #gives the length of each component of z
```

## Applying functions to lists: lapply()

Note that lapply() always assumes the input to be a list (e.g., a vector is coerced to a list form).  

```{r, label = ex2_lapply}
lapply(c(1:3), length)
```

## Applying functions to lists: sapply()

The output of lapply() is always a list. If you want to simplify the output (e.g., a vector): Use sapply() 


```{r, label = ex_sapply}
sapply(info, length)
sapply(SCR, length)
```


# Chapter 5: | Data frames

5.1  Creating Data Frames    

5.2  Matrix-Like Operations   

5.4  Applying functions to Data Frames   


## Introduction Data frames

5.1  Creating Data Frames

5.2  Matrix-Like Operations

*5.3  Merging Data Frames (self-study, important tool!)*

5.4  Applying functions to Data Frames



## Creating Data Frames
 
Data frames have properties of lists AND matrices... 

```{r, datafr1}
N <- 10
dfram <- data.frame(row.names = 1:N)
dfram$V1 <- c(1:9, NA)
dfram[[2]] <- c(NA, letters[1:9])
dfram[, 3] <- colors()[1:10]
dfram[, "V4"] <- 10:1
str(dfram)  # view structure of dfram
```

## Creating Data Frames
 
Data frames have properties of lists AND matrices... 

```{r, datafr2}
V1 <- c(1:9,NA)
V2 <- c(NA,letters[1:9])
V3 <- colors()[1:10]
V4 <- 10:1
dfram <- data.frame(V1,V2,V3, V4)
str(dfram) # view structure of dfram
```




## Extracting from Data Frames

Within in a 'local' data environment.

```{r, eval = TRUE}
with(data = dfram, expr = (V1 > 3) & (V4 > 3))
```

```{r, eval = TRUE}
subset(x = dfram, 
       subset = V1 > 3 & V4 > 3, 
       select = c(1,3,4)
       )
```



## List/Matrix-Like Operations

Which is List-like, which is Matrix-like? 

```{r, eval = FALSE}
class(dfram[["V1"]])
```

```{r, eval = FALSE}
class(dfram[, "V1", drop = FALSE])
```

```{r, eval = FALSE}
class(dfram[, "V1"])
```

```{r, eval = FALSE}
class(dfram["V1"])
```


## Matrix-Like Operations

When would (not) these code chunks work?

```{r, eval = FALSE}
apply(X = dfram, MARGIN = 2, FUN = mean, na.rm = TRUE)
```

```{r, eval = FALSE}
apply(X = dfram, MARGIN = 2, FUN = print)
```

```{r, eval = FALSE}
cbind(dfram, dfram[order(dfram$V4)[-4], ])
```

```{r, eval = FALSE}
rbind(dfram, dfram)
```


## Matrix-Like Operations

When would (not) these code chunks work?

```{r, eval = TRUE, message = FALSE}
apply(X = dfram, MARGIN = 2, FUN = mean, na.rm = TRUE)
```


## Matrix-Like Operations

When would (not) these code chunks work?

```{r, eval = TRUE, message = FALSE}
apply(X = dfram, MARGIN = 2, FUN = print)
```


## Matrix-Like Operations

When would (not) these code chunks work?

```{r, eval = TRUE, error = TRUE}
cbind(dfram, dfram[order(dfram$V4)[-4], ])
```



## Matrix-Like Operations

When would (not) these code chunks work?

```{r, eval = TRUE}
rbind(dfram, dfram)
```




## List-Like Operations

How about these implicit loops? 

```{r, eval = FALSE}
lapply(X = dfram, FUN = mean, na.rm = TRUE)
```

```{r, eval = TRUE}
sapply(X = dfram, FUN = mean, na.rm = TRUE)
```



## Recap of this afternoon


**Chapter 3: Matrices and arrays**   
  a. Unintended dimension reduction   
  b. Higher dimensional arrays   
  
**Chapter 4: Lists**   
  a. Creating lists and list properties   
  b. List indexing   
  c. Applying functions to lists   

**Chapter 5: Data Frames**   
  a. Creating Data Frames
  b. Matrix-Like Operations
  c. Applying functions to Data Frames


## Next Week

Chapters 6 - 7

See blackboard for the Peer-review assignment 2


