---
title: "Labmeeting on Animation"
author: "Maarten Kampert"
date: "18 Nov 2015"
output: html_document
---


## 1. If you did not manage yet,  get the lecture slides to knit!

## 2. Run the code given below in your console, and see what you are doing step by step.

```{r, label = rgl_plot, eval = FALSE}
library(rgl)

wadr <- "http://pub.math.leidenuniv.nl/~kampertmmd/OPSdat/red.green.csv"
redgreen <- read.csv(wadr)
attach(redgreen)

plot3d(c(-20, 45), c(-50, 80), c(0, 45), col = "white",
    xlab = "", ylab = "", zlab = "",    
    aspect = c(0.5, 1, 1.5), type = "n", lwd = 1, box = FALSE, axes = FALSE)
lines3d(c(-20, 45, 45, -20, -20), c(-50, -50, 80, 80, -50), c(0, 0, 0, 0, 0))
lines3d(c(-20, 45, 45, -20, -20), c(-50, -50, 80, 80, -50), c(45, 45, 45, 45, 45))
lines3d(c(-20, -20), c(-50, -50), c(0, 45))
lines3d(c(-20, -20), c(80, 80), c(0, 45))
lines3d(c(45, 45), c(-50, -50), c(0, 45))
lines3d(c(45, 45), c(80, 80), c(0, 45))


userMatrixRG<- matrix( c( 
    0.9, 0.45,0 , 0, 
    0, 0, 1,  0, 
    0.45, -0.9, 0, 0, 
    0, 0, 0, 1),
    4, 4, byrow = TRUE)


par3d(userMatrix = userMatrixRG)
 
plot3d(X, Y, T, type = "s", size = 1,col = Colour, add = TRUE)

plot3d(X, Y, T, type="l", add = TRUE)

play3d(spin3d())
```
To see for what purposes such code has been used, read from page 19 (Aim and Visualization) of the Statistical Science Master thesis of Helena van Eijk: http://www.math.leidenuniv.nl/scripties/MastervanEijk.pdf


## 3. Let's animate! FFmpeg.

With the help of an `FFmpeg` software program you can make your own videos in `R` using the `animate` package. In this exercise we hope to help you to make a video animation in R.

**For Windows:**

Download either the Static version of FFmpeg.exe in [64-bits](http://ffmpeg.zeranoe.com/builds/win64/static/ffmpeg-20151117-git-e9aea6d-win64-static.7z) or [32-bits](http://ffmpeg.zeranoe.com/builds/win32/static/ffmpeg-20151117-git-e9aea6d-win32-static.7z) from [ffmpeg.zeranoe.com/builds/#](ffmpeg.zeranoe.com/builds/#).

Open the zip file and move or copy the FFmpeg.exe file to the same directory as the directory in which you can find the .Rmd file with which you have knitted this .html file.

Unzip the file and find the directory name in which the `ffmpeg.exe` is to be found. Make your own `\ffmpeg\bin` directory 
in which you can find the (or a copy of) the `ffmpeg.exe`, e.g. `C:\ffmpeg\bin`. Add this directory to yout path by opening cmd.exe, and type `set path=%PATH%;C:\ffmpeg\bin`. **BE VERY CAREFUL NOT TO ACCIDENTILY DELETE YOUR PATH.**

Also take a look on [Stackoverflow](http://stackoverflow.com/questions/16481252/how-to-use-fig-show-animate-in-knitr-without-administrator-privileges) where Yihui refers you to [this website on how to set a PATH](http://www.computerhope.com/issues/ch000549.htm)


**For Mac OS X:**

Download FFmpeg for OS X from the following website: 

http://www.osxexperts.net/ffmpeg/ffmpegexperts.html

Run the downloaded `ffmpeg.....dmg` file, to create the FFmpeg file. Move or copy the FFmpeg file to the same directory as the directory in which you can find the .Rmd file with which you have knitted this .html file.

```{r, eval = FALSE}
if(Sys.which("ffmpeg") == ""){
	Sys.setenv(
	  PATH = paste(Sys.getenv("PATH"),getwd(),sep=":")
	)
}
```

Another and more permanent way would be to type or copy the following output to the code in your terminal on the mac:

```{r}
cat('sudo cp ', getwd(),'/ffmpeg /usr/local/bin/', sep = "")
```

**For Linux, Solaris:**

Probably you can manage yourself without instructions. If not, contact Maarten for this assignment.



**Make a clock**

Reproduce the clock as is shown on 
[stackoverflow](http://stackoverflow.com/questions/12038893/combining-r-markdown-and-animation-package)



```{r clock, fig.width=7, fig.height=6, fig.show='animate'}
par(mar = rep(3, 4))
for (i in seq(pi/2, -4/3 * pi, length = 12)) {
    plot(0, 0, pch = 20, ann = FALSE, axes = FALSE)
    arrows(0, 0, cos(i), sin(i))
    axis(1, 0, "VI"); axis(2, 0, "IX")
    axis(3, 0, "XII"); axis(4, 0, "III");
}
```

or try to reproduce your own video of Pacman.

```{r}
```


## 3. Google Visualization.

Try to reproduce the Google Visualizaton from the lecture slides about the members of the Statistical Society of the Netherlands.

To do so you need to transform the data object `VVS` in the right format, first.

```{r}
a = 56; b = 13; ib = rep(NA, 69- a-b)
VVS <- list(
  rnr = c(1:a, ib, (69-b + 1):69),
	Year = 1947:2015,
  # Exploitatie / Saldo  
  Rslt = c(NA,-(420-235)/2.20371, 617.03/2.20371, 78.37/2.20371, -(437.17/2.20371), 
           1517/2.20371, 2201.94/2.20371, 2276.81/2.20371, -3118.50/2.20371, NA,
           (870.53 - 71.73)/2.20371, (1623.16-819.17)/2.20371, 3584.02/2.20371, -1922.61/2.20371, 
           -(3816.39-357.12)/2.20371, -(4876.68)/2.20371, 2712.90/2.20371, -4814.39/2.20371, 
           1349.65/2.20371, rep(NA, 6),  -7917.95/2.20371, -20413.31/2.20371, -15115/2.20371, 
           30264/2.2037, 5958/2.20371, 19660/2.23071, 8154/2.20371, -4098/2.20371, 8088/2.20371, 
           rep(NA,3), 23335.91/2.20371, 63737.45/2.20371, 41235.95/2.20371, -10566/2.20371, 
           -60531/2.20371, -15845/2.20371, (7902)/2.20371, -9417/2.20371, 25982/2.20371, 
           7071/2.20371, 7787/2.20371, -6641/2.20371, 5555/2.20371, -1340/2.20371, -27201/2.20371,
           -6865/2.20371, 35227/2.20371, 4631, -6492, 561, -5101, -710, 11328, -8871, -7010, 72835, -17567, -23494,-7152, 5390, 13360, 
           NA ),
  # Nr of members	
	N = c(NA, 166,182,210,268,300,386, 426, 460, 534, 625, 694,698, 718, 754, 788, 834, 878, 929, 937,
        976, 1070, 1134, 1150, 1175, 1157, 1155, 1137, 1193, 1185, 1208, 1348, 1445, 1444, 1343,
        1316, 1249, 1251, 1286, 1281, 1299, 1301, 1423, 1475, 1480, 1445, 1430, 1402, 1391, 1354,
        1318, 1289, 1279, 1230, 1161, 1146, 1099, 1025,966,907,856,832,796,769,765,790,865,880,877),
#  / No section
#  n.NO = c(rep(NA, 12), 96, 95, 86, 77, rep(NA, 40), rep(NA, b) ),
#  / Bedrijfssectie
 	INDS = c(NA, NA,38,54,75,86,138,148, 160, 185,273, 313, 313, 312, 317, 324, 331, 360
            , 367, 370, 368, 342, 330, 330, 317, 280, 265, 270, 264, 230, 222, 212, 211,
            199, 184, 176, 162, 161, 154, 149, 153, 165, 198, 195, 216, 278, 291, 288, 279, 
            276, 280, 265, 265, 250, 231, 217, 204, 188, 169, 159, 152, 156, 135, 128, 117,
            107, 106, 100, NA),
# IS,MBS,MATH,OFF,SOC,ES,OR,AGR,SSP
#  / Medisch Biologische sectie / BIOSetric Section
	BIOS = c(rep(NA, 6), 63, 65, NA, NA,NA,NA, 122, 109, 121, 126, 132, 123, 157, 158,
            156, 165, 160, 163, 159, 148, 140, 146, 150, 137, 135, 133, 131, 141, 143,
            134, 138, 148, 141, 148, 150, 159, 179, 184, 181, 276, 296, 300, 298, 303,
    # called the BIOSetric section...
            366, 359, 357, 351, 333, 328, 325, 309, 287, 276, 262, 262, 243, 222, 225, 
    228, 235, 227, 227),
# / Sectie Beschrijvende (official stat.) statistiek / van 1964: SWS? / 
  OFFS = c(rep(NA, 6), 74, 77, 73, NA, NA, NA, 96, 96, 100, 102, 104, 99, rep(NA, 51)),
# / Sectie Mathematische Statistiek
	MATH = c(rep(NA, 6),69,73, NA,NA, NA,NA, 139, 136, 140, 150, 166, 158, 171, 188,
           201, 219, 241, 254, 277, 307, 320, 338, 354, 354, 381, 400, 415, 406, 408,
           399, 376, 366, 351, 346, 349, 331, 359, 355, 377, 441, 451, 452, 452, 428,
           415, 401, 370, 356, 340, 342, 331, 309, 295, 271, 265, 265, 244, 218, 227, 226, 
           232, 226, 237),
# / Sectie Economische Statistiek
	ECNS = c(rep(NaN, 9),NA,57,123, 177, 184, 200, 211, 232, 247, 262, 264, 253, 244,
           250, 247, 248, 247, 239, 252, 258, 231, 236, 236, 236, 226, 221, 217, 193, 
           183, 173, 169, 170, 173, 199, 192, 197, 269, 270, 270, 266, 254, 251, 240,
           240, 223, 212, 203, 187, 172, 157, 146, 140, 144, 128, 124, 117, 110,111, 104, 
           93),
# / Sectie Operational (Operations) Research / 97. NGB
  OPRS = c(rep(NaN, 12), 91, 97, 115, 151, 161, 194, 256, 277, 282, 292, 316, 335, 386,
           414, 433, 458, 471, 455, 460, 467, 536, 509, 445, 439, 411, 405, 380, 366, 375,
           369, 426, 433, 457, 528, 510, 503, 493, 477, 462, 456, 438, 404, 371, 374, 349,
           316, 289, 279, 265, 283, 273, 240, 244, 253, 261, 299, 269),
# Landbouwkundige Sectie
  AGRI = c(rep(NaN, 16), 15, 19, 31, 33, 46, 50, 56, 61, 62, 59, 60, 62, 60, 64, 66,
           64, 68, 68, 71, 68, 66, 67, 56, 56, 63, 67, 76, 75, 80, 178, 191, 190, 189,
           188, rep(NA,6), rep(NA, b) ),
  SOCS = c(rep(NA, 6 + 12), #SWS begins....
         83, 116,
         121, 120, 112, 108, 109, 103, 97, 102, 112, 110, 124, 278, 334, 355, 320,
         321, 303, 281, 256, 252, 249, 251, 280, 277, 283, 371, 375, 375, 376, 368,
         364, 351, 346, 343, 337, 325, 321, 286, 268, 253, 240, 229, 211, 184, 194, 
         201, 222, 215, 208),
# Sectie Statistische Programmatuur (1982 for real, in 1968 werkgroep/contactgroep Statistiek en Computer)
  SSFP = c(rep(NaN, 36), 86, 105, 127, 138, 153, 182, 217, 229, 227, 313, 323, 320, 319, 309,
           301, 288, 273, 261, 243, 231, 220, 202, 186, 163, 155, 148, 136, rep(NA, 6) )
)
```

```{r}

```

Eventually use the following command, as was used in the lecture slides.
```{r}

```

