---
output: pdf_document
---

<!-- To knit this .Rmd file you should have installed the nycflights13 package -->

# Assignment Week 1:

# Working with Vectors and Filtering Indices

\vskip 1cm

- Yi He

- s1684698





NOTES: 

make a reproducable report of this assignment according to report requirements we have set in this course. See blackboard for more information.    

Set al R code chunks to eval = TRUE.   

When you have difficulties making this assignment it is advisable to take the first six lessons of the R programming course of the R package 'swirl'.


\newpage 

### 1. Filtering and indices

First load the package `nycflights13` using the following code chunk.

```{r, label = preps, message = TRUE}
library(nycflights13)
```

Before you start with the assignment find the `nycflights13` package documentation using `help(package = 'nycflights13')` and take a look at the documentation of the `flights` objects. 

Having read the documentation, fill out here your code to see the class and the dimensions of the `flights` object.
```{r, label = classdim}
class(flights) 
dim(flights) 
```

In this data object we would like you to create a logical vector in each code chunk such that we can find all flights...   

...to SFO or OAK
```{r label = flt1, eval = TRUE}
flt.1 <- c(flights$dest == 'SFO' | flights$dest == 'OAK') 
```

There are `r sum(flt.1, na.rm = TRUE)` flights to SFO or OAK.

... delayed by more than an hour   
```{r label = flt2, eval = TRUE}
flt.2 <- c(flights$dep_delay > 60) 
# Not sure to use dep_delay or arr_delay
```
There are `r sum(flt.2, na.rm = TRUE)` flights delayed by more than an hour.

... that departed between midnight and five am 
```{r label = flt3, eval = TRUE}
flt.3 <- c(flights$dep_time < 500 & flights$dep_time > 0) 
# Exclude the end points of the interval.
# Use table() to compare flights$hour and flights$dep_time.
# flights$hour has 29 observations on 24.
# flights$dep_time shows that they were exactly on 2400.
```
There are `r sum(flt.3, na.rm = TRUE)` flights that departed between midnight and five am.

... for which the arrival delay was more than twice the departure delay   

```{r label = flt4, eval = TRUE}
flt.4 <- c(flights$arr_delay > 2 * flights$dep_delay)  
```

There are `r sum(flt.4, na.rm = TRUE)` flights for which the arrival delay was more than twice the departure delay.

\newpage

### 2. Creating Variables

First run the following code to create a data frame `data.set` with no variables:

```{r, label = data.set, eval = TRUE}
set.seed(20151029) 
indx.flights <- sample(1:nrow(flights), 1e3) 
N <- 1e3 
data.set <- data.frame(row.names = 1:N) 
```



Now, we would like to add the following variables. 

- `x1`: numeric vector with components \(N/2,N/2 -1, N/2 - 2,\ldots,2,1,1,2 \ldots,N/2 - 2, N/2 - 1,N /2\).

- `x2`: a logical vector of lenght N. Extract the carrier information out of the `flights` data for the rownumbers which can be found in `indx.flights`. For each element that has 'US','UA' or 'AA' as a carrier will obtain a `TRUE`, all others `FALSE`.

- `x3`: a 'bad weather' indication variable. Draw \(N\) observations that come from an exponential distribution with a rate parameter equal to \(1/10\). Then, round these observations to the closest integer (use the functions `rexp` and `round`).

- `e`: a vector of \(N\) elements that come from a standard normal i.i.d.

- `y`: our own synthetic or simulated outcome variable representing arrival delay: \(y_i = 2 + 0.1*x_{i2} + 2*x_{i3} + 1*x_{i2}*x_{i3} + e_i\).



```{r, label = vars, eval = TRUE}
data.set$x1 <- abs((N/2):-(N/2))[-(N/2 + 1)] 
data.set$x2 <- flights$carrier[indx.flights]%in%c("AA","UA","US") 
data.set$x3 <- round(rexp(N, 1/10), 0) 
data.set$e <- rnorm(N) 
data.set$y <- with(data.set, 2 + 0.1*x2 + 2*x3 + 1*x2*x3 + e) 
```










### 3. Log-Likelihood Ratio Statistic.

We have the folliwing linear regression model:

\(y_i = \beta_0 + \beta_1*x_{i2} + \beta_2*x_{i3} + \beta_3*x_{i2}*x_{i3} + \epsilon_i = \hat{y}_i + \epsilon_i \).


**3a.** We can obtain estimates of the values of the \(\beta\)'s and \(sigma\) using the function `lm()`;

```{r label = rgr1}
lm.rgr1 <- with(data.set, lm(y ~ x2*x3)) 
paste0("hat.b",0:3, " = " , round(lm.rgr1$coef, 2)) 
sigma.h <- sd(lm.rgr1$residuals) 
paste0("sigma.h = ", round(sigma.h, 2)) 
```
Using these estimates, what is the value of the log likelihood of our model? 

Hint: use the function `sum` and know log-likelihood can be calculated as follows

\[
 \log \left( 
 \prod^N_{i = 1} \frac{1}{\hat{\sigma} \sqrt{2 \pi}} \exp^{ 
    -\frac{1}{2} \left( \frac{ y_i - \hat{y}_i) }{ \hat{\sigma} } \right)^2} 
    \right)
\]

```{r, label = L_h1, eval = TRUE}
L.h1 <-  (N * log(1 / (sigma.h * sqrt(2 * pi))) 
          - sum((lm.rgr1$residuals) ^ 2) / (2 * sigma.h ^ 2)) 
# Not sure of the google R style.
# Should we put spaces everytime we use + - * / ^ ?
# It looks so strange.
```

The value of the log-likelihood (\(l_1\)) of data set equals `r L.h1`.


**3b.** Suppose we would like to add the x1 variable in our model as follows:

\[y_i = \beta_0 + \beta_1*x_{i1} + \beta_2*x_{i2} + \beta_3*x_{i3} + \beta_4*x_{i2}*x_{i3} +\epsilon_i = \hat{y}_i + \epsilon_i \].

Using the estimates from the following code chunk:

```{r label = rgr2}
lm.rgr2 <- with(data.set, lm(y ~ x1 + x2*x3)) 
paste0("hat.b",0:4, " = " , round(lm.rgr2$coef, 2)) 
sigma.h <- sd(lm.rgr2$residuals) 
paste0("sigma.h = ", round(sigma.h, 2)) 
```

What is the value of the log-likelihood?

```{r, label = L_h2, eval = TRUE}
L.h2 <- (N * log(1 / (sigma.h * sqrt(2 * pi)))  
         - sum((lm.rgr2$residuals) ^ 2) / (2 * sigma.h ^ 2)) 
```

**3c.** Calculate the log-likelihood ratio of our first model against the second model:

```{r, label = LRT, eval = TRUE}
LRT <- L.h1 / L.h2
```

When the log-likelihood ratio exceeds the value of \(\approx 3.84\), then we find a significant contribution for \(\alpha = 0.05\) for our extra `x1` variable. Here the value of the log-likelihood ratio equals `r round(LRT, 2)` and corresponds to a p-value of approximately `r round(pchisq(LRT, df = 1, lower.tail = FALSE), 4)`. Hence, as expected, adding the `x1` variable to the model does not improve model fit.
