---
title: "I Just Ran Two Million Regressions 4-2"
author: "Wang ZeXian"
date: "2016年5月12日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Xavier X. Sala-I-Martin: I Just Ran Two Million Regression (AER, 1997)          
Growth Convergence of 72 countries with 41 variables            
Principal Components Analysis           
# setwd("C:/Course16/WISE2016/R")
http://web.pdx.edu/~crkl/WISE2016/data/FLS-data.csv
```{r}
#growth <- read.csv("C:\\Users\\44180\\Documents\\Surface-workandstudy\\soe\\bigdata\\data of case4\\FLS-data.csv")
growth <- read.csv("D:\\PC-workandstudy\\soe\\bigdata\\data of case4\\FLS-data.csv")
summary(growth)
```

# principal components analysis主成份分析算法
```{r}
#working with data matrices
X <- as.matrix(growth[,-1])
Y <- as.matrix(growth[,1])

# PCA on X
# compute correlation matrix of X
R <- cor(X)
# find eigenvalues and eigenvectors of R
EV <- eigen(R)
str(EV)
# EV$Values are the variances of PC特征值是主成份的方差
# EV$Vectors are the loadings特征向量是主成份的载荷
PCsd <- sqrt(EV$values)
PCloadings <- EV$vectors
rownames(PCloadings) <- colnames(X)
#主成份得分
PCscores <- X %*% PCloadings
print(PCsd)
print(PCloadings)
print(PCscores)
plot(PCscores[,1], PCscores[,2], xlab="PCA 1", ylab="PCA 2",
     type="n", xlim=c(min(PCscores[,1:2]), max(PCscores[,1:2])),
     ylim=c(min(PCscores[,1:2]), max(PCscores[,1:2])))
arrows(0,0,PCloadings[,1]*3,PCloadings[,2]*3, length=0.1,
       angle=20, col="red")
text(PCloadings[,1]*3*1.2,PCloadings[,2]*3*1.2,
     rownames(PCloadings), col="red", cex=0.7)
text(PCscores[,1],PCscores[,2], rownames(PCscores), col="blue",
     cex=0.7)
```

# principal components analysis主成份分析(基于函数)
# using princomp, prcomp, factanal packages
# data will be scaled in computation数据会被自动标准化
```{r}
# PCA on X
pca <- princomp(X, cor=TRUE)
summary(pca)
pca$scores
pca$loadings
#第一主成份的各个国家(样本)得分
plot(pca$scores[,1])
barplot(pca$scores[,1])
#用第1-12个主成份的得分代表各个样本进行回归
reg1 <- lm(Y~pca$scores[,1:12])
summary(reg1)
#用第21-41个主成份的得分代表各个样本进行回归
reg2 <- lm(Y~pca$scores[,21:41])
summary(reg2)
```

# using prcomp package主成份分析(基于包)
# data must be scaled, otherwise set center and scale.
```{r}
prc <- prcomp(X)
summary(prc)
#特征值根号
prc$sdev
#载荷
prc$rotation  # loadings
#得分
prc$x         # scores
```

#using factanal package(基于包-因子分析模型)
```{r}
fa <- factanal(X, factors=12, scores="regression")
summary(fa)
fa$loadings
fa$scores

reg3 <- lm(Y~fa$scores)
summary(reg3)
```
