---
title: "Credit Scoring and Default Prediction 5-2"
author: "Wang ZeXian"
date: "2016年5月12日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

AmEx Credit Scoring and DEFAULT
DEFAULT|CARDHLDR ~ AGE+INCOME+OWNRENT+SELFEMPL+ACADMOS, used by Greene's lecture
```{r}
# setwd("C:/Course16/WISE2016/R")
#AmEx<-read.csv("C:\\Users\\44180\\Documents\\Surface-workandstudy\\soe\\bigdata\\data of case5\\AmEx.csv")
 AmEx<-read.csv("D:\\PC-workandstudy\\soe\\bigdata\\data of case5\\AmEx.csv")
# exclude<-c(1,2,"BANKBOTH","CREDGAS","UNEMP","DRUGSTOR")
# DEFAULT|CARDHLDR=1(default定义为1)
AmEx1<-subset(AmEx,CARDHLDR==1,select=-c(1,6,12,17,39:56))
Y<-as.matrix(AmEx1[,"DEFAULT"])
Xs<-as.matrix(AmEx1[,-1])
```

for qualitative choice model, package lars can not used,it designed for qualitative regression                  
for LAR, LASSO, and CV-based selection(we can use glmnet package)
```{r}
library(glmnet)
# for linear regression, glm with family=gaussion
# for logit regression, glm with family=binomial
# Default equation conditional to Cardholder
# Ridge Regression,alpha=0 means Ridge Regression
logit1<-glmnet(Xs,Y,family="binomial",alpha=0)
logit1
#L1 Norm L2惩罚函数=\sum _{i=1}^{p} (\beta_i)^{2}
plot(logit1)
#所有参数同时坍缩到0,无法用于model selection
plot(logit1,xvar="lambda")
coef(logit1,s=0.1)  # s=lambda selected(这里随便选的)

# LASSO for variables selection,alpha=1 means LASSO
logit2<-glmnet(Xs,Y,family="binomial",alpha=1)
logit2
#L1 Norm L1惩罚函数=\sum _{i=1}^{p} |\beta_i|
#参数有先后收缩到0
plot(logit2)
#可以看到变量参数坍缩到0的先后顺序
plot(logit2,xvar="lambda")

# cross validation to pick lambda: 10-fold CV
#数据较大,可以分10块(随机分块),重复cv好多次,每次选出一个最小lambda,再从最小lambda中选出中位数作为最优lambda
#要set seed才能保证每次都一样
set.seed(12450)
#岭回归不能选择变量,只做LASSO
logit2.cv <- cv.glmnet(Xs,Y,family="binomial",alpha=1)
#纵轴:错误分类率(miss classification rate)
plot(logit2.cv)
# pick best lambda标线的是最优,seed不同结果不同
best <- logit2.cv$lambda.min
best   # check log(best)
log(best)
#coef <- predict(logit2,s=best,type="coef")
logit2.coef<-coef(logit2,s=best)
logit2.coef
logit2.coef[logit2.coef!=0]

# repeated CV method may be used for logit2 model (not run)
# this can take a long time tor run
# 重复cv100次,每次选出一个最优lambda,从全部lambda中选出中位数作为selected
cv_vector<-rep(0,100)
for(i in 1:100) {
  logit2.cv <- cv.glmnet(Xs, Y, family="binomial", alpha=1)
  cv_vector[i]=logit2.cv$lambda.min
}
median(cv_vector)  # find median value of #lambda.min
coef(logit2,s=median(cv_vector))  # s=lambda selected 

