---
title: "Chinese Yuan and Stock Market 6-4"
author: "Elara"
date: "2016年5月14日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# China Stock Market Index: Shanghai Common Stock
Time Series Exponential Smoothing using forecast package
```{r}
# Using Quandl API to search and retrieve data series
# install.packages("Quandl")
library(Quandl)
# Quandl.auth("4UqGASL7CigqHzpfdnNw")
# SP500<-Quandl(code="YAHOO/INDEX_GSPC",type="xts") 
# xts or zoo type should be used for daily high frequency data
# frequency conversion for xts type: apply.xxx
# xxx=yearly, quartertly, monthly, weekly
#SSEC<-Quandl(code="YAHOO/INDEX_SSEC",type="xts") 
# daily price index: irregularly-spaced time series
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case6\\SP500 and SSEC.RData")
data<-SSEC[,"Adjusted Close"]
summary(data)
plot(data)
```
# monthly average: regularly-space time series
```{r}
mdata<-ts(apply.monthly(data,mean),start=c(1997,7),frequency=12)
plot(mdata)
```
# daily returns in percent
```{r}
rd<-na.omit(100*diff(log(data)))
rm<-ts(apply.monthly(rd,sum),start=c(1997,7),frequency=12)
plot(rm,main="Monthly Log Returns",ylab="Percent")

# rv<-ts(apply.monthly(rd^2,sum),start=c(1997,7),frequency=12)
```
# select variable, setup train and test data    
```{r}
Y<-rm
Y.train<-window(Y,end=c(2014,12))
Y.test<-window(Y,start=c(2015,1))
```
# ETS Analysis and Forecasts    
```{r}
library(forecast)
# model=(level, trend, seasonal)
# ets1<-ets(Y.train,model="ZZZ")  # automatic model(ZZZ automatic using machine learning to generate a model)
#level type ("A", "M" or "Z");trend type ("N","A","M" or "Z"); season type ("N","A","M" or "Z").  "N"=none, "A"=additive, "M"=multiplicative and "Z"=automatically selected. 
#example, "ANN" is simple exponential smoothing with additive errors, "MAM" is multiplicative Holt-Winters' method with multiplicative errors, and so on.
ets1<-ets(Y.train,model="AAA")
ets1
plot(ets1)
```

# h-step ahead forecasts
#没有信息,直接预测
```{r}
for1<-forecast(ets1,h=30)
for1
plot(for1)
lines(Y.test)
accuracy(for1,Y.test)
```

# time series cross validation
time series has autocorrelation,so we cant devide it into n\<k fold and we cant select it randomly.But we can devide data into k-flor randomly,only use 1-step ahead        
# 1-step ahead forecasts without re-estimation          
# 每次只往前预测一步,用的模型是train生成的模型          
```{r}
ets1a<-ets(Y,model=ets1)  # fit entire data with estimated model,it is Y ,not Y.train(using the traning data to fit entire data)
for1a<-window(fitted(ets1a),start=c(2015,1)) # extract the forecasts
for1a
plot(for1)
lines(Y.test)
lines(for1a,col="red")#blue:30-step ahead forecast,red:1-step ahead (using the same model)
accuracy(for1a,Y.test)
```

# a function to compute forecast error statistics
```{r}
predict.error<-function(x,p) {
  data<-ts.intersect(x,p)
  x<-data[,"x"]
  p<-data[,"p"]
  e<-x-p                           # prediction error
  mx<-mean(x)
  mp<-mean(p)
  sx<-sqrt(mean((x-mx)^2))
  sp<-sqrt(mean((p-mp)^2))
  r<-mean((x-mx)*(p-mp))/(sx*sp)
  # r<-cor(x,p)
  MSE<-round(mean(e^2),4)                   # mean squared error
  # results list
  r2<-round(r^2,4)
  ME<-round(mean(e),4)                      # mean error
  MAE<-round(mean(abs(e)),4)                # mean absolute error
  MAPE<-round(100*mean(abs(e/x)),4)         # mean absolute error
  RMSE<-round(sqrt(MSE),4)                  # root mean squared error
  RMSPE<-round(100*sqrt(mean((e/x)^2)),4)   # root mean squared error
  # MSE Decomposition in percent
  Um<-round(100*((mx-mp)^2)/MSE,4)
  Us<-round(100*((sx-sp)^2)/MSE,4)
  Uc<-round(100*(2*(1-r)*sp*sx)/MSE,4)
  Ur<-round(100*((sp-r*sx)^2)/MSE,4)
  Ud<-round(100*((1-r^2)*sx^2)/MSE,4)
  results<-c(ME,MAE,MAPE,MSE,RMSE,RMSPE,Um,Us,Uc,Ur,Ud,r2)
  names(results)<-c("ME","MAE","MAPE","MSE","RMSE","RMSPE","Um%","Us%","Uc%","Ur%","Ud%","r2")
  return(results)
}
```

# prediction error statistics
```{r}
predict.error(Y.test,for1$mean)
predict.error(Y.test,for1a)
```