---
title: "Chinese Yuan and Stock Market 6-5"
author: "Wang ZeXian"
date: "2016年5月15日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# China Stock Market Index: Shanghai Common Stock
State Space Time Series Analysis using dlm package

# Using Quandl API to search and retrieve data series
```{r}
# install.packages("Quandl")
library(Quandl)
# Quandl.auth("4UqGASL7CigqHzpfdnNw")
# SP500<-Quandl(code="YAHOO/INDEX_GSPC",type="xts") 
# xts or zoo type should be used for daily high frequency data
# frequency conversion for xts type: apply.xxx
# xxx=yearly, quartertly, monthly, weekly
#SSEC<-Quandl(code="YAHOO/INDEX_SSEC",type="xts") 
# daily price index: irregularly-spaced time series
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case6\\SP500 and SSEC.RData")
data<-SSEC[,"Adjusted Close"]
summary(data)
plot(data)
```             

# select a variable for analysis and DLM
DLM page 4 in [TSAR-2.pdf](C:\Users\44180\Documents\R-in-SOE\R\big data\case6\TSAR-2.pdf)
```{r}
Y<-log(data)

level0<-mean(Y) #mean of Y
slope0<-mean(diff(Y),na.rm=T) #mean of diff Y

library(dlm)
#输入params,参数
# local level model
dlm1<-function(params) {
  dlmModPoly(order=1,dV=exp(params[1]),dW=exp(params[2]))
        # exp:make sure it is positive
        # it is a linear model
        # order=1:level,dv--v_t,dw--w_t
}
# constant level: dW=0, restricted
# w_t的差分为0,即为常数
dlm11<-function(params) {
  dlmModPoly(order=1,dV=exp(params[1]),dW=0)
}

# local trend model
#trend出现在state equation中,,observation equation中仍然只有v_t,state equation出现2个w_t
dlm2<-function(params) {
  dlmModPoly(order=2,dV=exp(params[1]),dW=exp(params[2:3]),m0=c(level0,slope0))
        # dv 1 equation,dw 2 equations,m0 is given
        #m0是初始状态分布(initail state distribution)
}
# local trend with constant level: dW[2]=0 restricted
#w的第二个参数差分为0,trend项不变
dlm2<-function(params) {
  dlmModPoly(order=2,dV=exp(params[1]),dW=c(0,0),m0=c(level0,slope0))
        #第一项差分也为0?
}

# local trend ARMA model
dlm3<-function(params) {
  dlmModPoly(order=2,dV=exp(params[1]),dW=exp(params[2:3]),m0=c(level0,slope0)) +  
    dlmModARMA(ar=params[5],ma=params[6],sigma2=exp(params[4]))
}
```

# ML Estimation 
```{r}
dlm.fit<-dlmMLE(Y,c(1,1),dlm1)#(data,initial,model)
#dlm.fit<-dlmMLE(Y,c(1,1,1),dlm2)
#dlm.fit<-dlmMLE(Y,c(1,1,1,1,0,0),dlm3)
dlm.fit# need to cheak converage =0?if not the solution is not realiable
dlm.model<-dlm1(dlm.fit$par)
#dlm.model<-dlm2(dlm.fit$par)
#dlm.model<-dlm3(dlm.fit$par)
dlm.model
dlm.model$V  # V(dlm1.model)
dlm.model$W  # W(dlm1.model)
dlm.filtered<-dlmFilter(Y,dlm.model)# 2 step first estimation,second, update
dlm.smoothed<-dlmSmooth(dlm.filtered)# 
dlm.resid<-residuals(dlm.filtered)$res
dlm.fm<-dropFirst(dlm.filtered$m)  # filtered mu
dlm.sm<-dropFirst(dlm.smoothed$s)  # smoothed mu
ts.plot(dlm.sm)
#ts.plot(dlm.sm[,1])
#ts.plot(dlm.sm[,2])
