---
title: "Chinese Yuan and Stock Market 6-2"
author: "Elara"
date: "2016年5月14日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# China Stock Market Index: Shanghai Common Stock
# Structural Time Series (StrucTS function)
#
# Using Quandl API to search and retrieve data series
# install.packages("Quandl")
library(Quandl)
# SP500<-Quandl(code="YAHOO/INDEX_GSPC",type="xts") 
# xts or zoo type should be used for daily high frequency data
# frequency conversion for xts type: apply.xxx
# xxx=yearly, quartertly, monthly, weekly
SSEC<-Quandl(code="YAHOO/INDEX_SSEC",type="xts") 
# daily price index: irregularly-spaced time series
data<-SSEC[,"Adjusted Close"]
summary(data)
plot(data)

# daily return as difference of log price index
```{r}
rd<-na.omit(100*diff(log(data)))
summary(rd)
plot(rd)
# monthly return is the sum of daily returns in a month
rm<-ts(apply.monthly(rd,sum),start=c(1997,7),frequency=12)
# monthly volatility is the sum of daily squared returns in a month
rv<-ts(apply.monthly(rd^2,sum),start=c(1997,7),frequency=12)
plot(rv)#risk level monthly ,ylab is wrong
Y<-rv
summary(Y)
```

# Exponential Smoothing: Holt-Winters Filtering指数平滑
```{r}
sts1<-decompose(Y)#把数据分解成趋势,季度,随机扰动
plot(sts1)
HoltWinters(Y)#求解HoltWinters指数平滑系数,ab是趋势,s是季度影响,平滑参数alpha:30%的过去影响,beta:没有斜率趋势=0
HoltWinters(Y,beta=F,gamma=F)#不含beta和gamma
```

# Exponential Smoothing: Structural TS(base)3种模型
```{r}
# local level model(no slope)
sts1a<-StructTS(Y,type="level",init=c(1,1))#init参数初始值
sts1a
plot(Y)
lines(sts1a$fitted,col="red")
# local linear trend model(allow trend)
sts1b<-StructTS(Y,type="trend",init=c(1,1,1))
sts1b#slope仍然是0
plot(Y)
lines(sts1b$fitted[,1],col="red")
# best structural model (linear trend and seasonal model)
sts1c<-StructTS(Y,type="BSM")#BSM=best structural model
sts1c#slope仍然是0,有season因素
plot(Y)
lines(sts1c$fitted[,1],col="red")

# plot of slope of the trend model
sts1b$fitted
plot(sts1b$fitted[,2])

# plot of slope and seasonal of BSM
sts1c$fitted
plot(sts1c$fitted[,2])
plot(sts1c$fitted[,3])
```