---
title: "Chinese Yuan and Stock Market 6-2"
author: "Elara"
date: "2016年5月14日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

# China Stock Market Index: Shanghai Common Stock
Structural Time Series (StrucTS function)
```{r}
#Using Quandl API to search and retrieve data series
# install.packages("Quandl")
library(Quandl)
# SP500<-Quandl(code="YAHOO/INDEX_GSPC",type="xts") 
# xts or zoo type should be used for daily high frequency data
# frequency conversion for xts type: apply.xxx
# xxx=yearly, quartertly, monthly, weekly
#SSEC<-Quandl(code="YAHOO/INDEX_SSEC",type="xts") 
# daily price index: irregularly-spaced time series
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case6\\SP500 and SSEC.RData")
data<-SSEC[,"Adjusted Close"]
summary(data)
plot(data)
```

# daily return as difference of log price index
```{r}
rd<-na.omit(100*diff(log(data)))# daily return rd
summary(rd)
plot(rd)
# monthly return is the sum of daily returns in a month
rm<-ts(apply.monthly(rd,sum),start=c(1997,7),frequency=12)# monthly return rm
# monthly volatility is the sum of daily squared returns in a month
rv<-ts(apply.monthly(rd^2,sum),start=c(1997,7),frequency=12)# monthly volatility rv
plot(rv)#risk level monthly ,ylab is incorrect
Y<-rv #keep rv
summary(Y)
```
# Exponential Smoothing
1. model:
$$
y_t=m_t+s_t+\epsilon_t \textrm{ or } ln(y_t)=m_t+s_t+\epsilon_t \\
trend:m_t=a_t+b_t(level+slope) \\
seasonal:s_t \\
random:\epsilon_t
$$              
2. EWMA(Simple Exponential Smoothing)   
case：$b_t=0$,$s_t=0$         
model：$y_{t+1}=a_t+b_t+s_t+\epsilon_t$
smoothing equation:     
$$
\begin{eqnarray}
a_t & = & \alpha y_t+(1-\alpha )a_{t-1},0<\alpha <1   \\
    & = & \alpha y_t+\alpha(1-\alpha)y_{t-1}+\alpha(1-\alpha)^{2}y_{t-2}+\dots \\
\end{eqnarray}
$$      
forecast equation:      
$$
\begin{eqnarray}
\hat{y}_{t+1|t} & = & a_t,t & = & 1,2,\dots,n \\
\hat{y}_{n+h|n} & = & a_n,h & = & 1,2,\dots,n \\
\end{eqnarray}
$$      
state space representation:      
$$
\begin{eqnarray}
y_t & = & a_{t-1}+e_t \textrm{  observation equation} \\
a_t & = & a_{t-1}+\alpha e_t \textrm{  state equation} \\
a_1 & = & y_1 \textrm{  initialization} \\
a_2 & = & a_1+\alpha(y_2-a_1) \textrm{  observation equation} \\
\dots & {} & {} \\
a_n & = & a_{n-1}+\alpha(y_n-a_{n-1}) \textrm{  } \hat{\alpha}=\min_{\alpha}arg\sum _{t=1}^{n}e_t^{2}
\end{eqnarray}
$$

3. Holt Exponential Smoothing(including trend)           
$s_t=0$         
page 8 in[TSAR-1.pdf](C:\Users\44180\Documents\R-in-SOE\R\big data\case6\TSAR-1.pdf)

4. Holt-Winters Exponential Smoothing(including trend and season)   
page 10 in[TSAR-1.pdf](C:\Users\44180\Documents\R-in-SOE\R\big data\case6\TSAR-1.pdf)

# Exponential Smoothing: Holt-Winters Filtering
```{r}
#Y=rv=monthly volatility,is the sum of daily squared returns in a month
sts1<-decompose(Y)#把数据分解成趋势,季度,随机扰动
plot(sts1)
HoltWinters(Y)#求解HoltWinters指数平滑系数,a是水平,b是趋势,s是季度影响,平滑参数alpha,beta(没有斜率趋势=0),gamma
HoltWinters(Y,beta=F,gamma=F)#设定不含beta和gamma
```

# Exponential Smoothing: Structural TS(base) 3 models
page 2 in [TSAR-2.pdf](C:\Users\44180\Documents\R-in-SOE\R\big data\case6\TSAR-2.pdf)
```{r}
# local level model(no slope)
sts1a<-StructTS(Y,type="level",init=c(1,1))#init初始vt,wt方差
sts1a
plot(Y)
lines(sts1a$fitted,col="red")
# local linear trend model(allow trend)
sts1b<-StructTS(Y,type="trend",init=c(1,1,1))#init初始扰动项方差
sts1b#slope仍然是0
plot(Y)
lines(sts1b$fitted[,1],col="red")
# best structural model (linear trend and seasonal model)
sts1c<-StructTS(Y,type="BSM")#BSM=best structural model
sts1c#slope仍然是0,有season因素
plot(Y)
lines(sts1c$fitted[,1],col="red")

# plot of slope of the trend model
sts1b$fitted
plot(sts1b$fitted[,2])

# plot of slope and seasonal of BSM
sts1c$fitted
plot(sts1c$fitted[,2])
plot(sts1c$fitted[,3])
```