---
title: "Chinese Yuan and Stock Market 6-8"
author: "Wang ZeXian"
date: "2016年5月15日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

# State Space Time Series Analysis using dlm package
Time varying parameters model (TVP)
China Stock Market Index (Shanghai Common Stock) and
Renminbi Exchange Rate (CNY/USD)
# Using Quandl API to search and retrieve data series
```{r}
# install.packages("Quandl")
library(Quandl)
# Quandl.auth("4UqGASL7CigqHzpfdnNw")
# SP500<-Quandl(code="YAHOO/INDEX_GSPC",type="xts") 
# xts or zoo type should be used for daily high frequency data
# frequency conversion for xts type: apply.xxx
# xxx=yearly, quartertly, monthly, weekly
#SSEC<-Quandl(code="YAHOO/INDEX_SSEC",type="xts") 
#CNY<-Quandl(code="BUNDESBANK/BBEX3_D_CNY_USD_CA_AC_000",type="xts") 
# daily price index: irregularly-spaced time series
# combine data with common rows
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case6\\SP500 and SSEC.RData")
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case6\\CNY.RData")
data<-merge(SSEC[,"Adjusted Close"],CNY,all=F)#only get the time both not na
colnames(data)<-c("SSEC","CNY")
# data<-subset(data,time(data)>c("2005-12-31"))
```
# static model
```{r}
Y<-log(data[,"SSEC"])
X1<-data[,"CNY"]
ols1<-lm(Y~X1)
summary(ols1)
```
# dlm
```{r}
library(dlm)
# time varying parameters model
tvp1<-function(params,data) {
  dlmModReg(X=data,dV=exp(params[1]),dW=exp(params[2:3]))
}

# ML Estimation 
tvp.fit<-dlmMLE(Y,c(1,1,1),data=X1,build=tvp1)
tvp.fit#it does not converage,so we cant do filter or smooth,may it be easy to restimate the data part into 2 part by the structual change
tvp.model<-tvp1(tvp.fit$par,X1)
tvp.model
tvp.model$V  # V(ssm1.model)
tvp.model$W  # W(ssm1.model)
tvp.filtered<-dlmFilter(Y,tvp.model)
tvp.smoothed<-dlmSmooth(tvp.filtered)
tvp.resid<-residuals(tvp.filtered)$res
tvp.sm<-dropFirst(tvp.smoothed$s)
ts.plot(tvp.sm)
ts.plot(tvp.sm[,2]) # TVP parameter of X1

acf(tvp.resid)
pacf(tvp.resid)

