---
title: "Chinese Yuan and Stock Market 6-1"
author: "Elara"
date: "2016年5月14日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

China Stock Market Index: Shanghai Common Stock(A股)
Time Series Data Visualization
#数据获取
```{r}
# Using Quandl API to search and retrieve data series
# install.packages("Quandl")
library(Quandl)
# Quandl.api_key("4UqGASL7CigqHzpfdnNw")
# SP500<-Quandl(code="YAHOO/INDEX_GSPC",type="xts") 
# xts:regular format
# xts or zoo type should be used for daily high frequency data
# frequency conversion for xts type: apply.xxx
# xxx=yearly, quartertly, monthly, weekly
# SSEC<-Quandl(code="YAHOO/INDEX_SSEC",type="xts") 
# daily price index: irregularly-spaced time series
# the following is for saving to csv and retriving from it
# SSEC_raw<-Quandl(code="YAHOO/INDEX_SSEC",type="raw",order="asc") 
# write.csv(SSEC,file="C:/Course16/WISE2016/data/SSEC.csv",row.names=F)
# SSEC1<-read.csv(file="C:/Course16/WISE2016/data/SSEC.csv",stringsAsFactors=F)
# SSEC1<-xts(SSEC1,order.by=as.Date(SSEC1$Date))
# data<-SSEC1[,"Adjusted.Close"]
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case6\\SP500 and SSEC.RData")
#Volume数据开始为0是因为没数据
```

#数据转换
```{r}
data<-SSEC[,"Adjusted Close"]
class(data)
summary(data)
plot(data)
mdata<-ts(apply.monthly(data,mean),start=c(1997,7),frequency=12)
# monthly average: regularly-space time series把日值转化成月度均值
# ts:time series
# apply.monthly(data,function):solve the function of the data monthly
# start=c(yyyy,m)
# frequency=the number of observations per unit of time.(only a setting value,not influence the mdata)
class(mdata)
head(mdata)
plot(mdata)
```

#月度趋势
```{r}
# install.packages("forecast")
library(forecast)
monthplot(mdata)#每张小图是97-16年某月的月平均指数,横线是不同年份在该月指数的均值
seasonplot(mdata,s=12,year.labels=T,col=rainbow(12))#每条线是某一年各月均值
lag.plot(mdata,12)#滞后相关图
```

#daily return as difference of log price index
#日趋势
```{r}
#日收益率rd=100*diff(log(data))=100*ln(data@t)-ln(data@t-1)=100*ln(data@t/data@t-1)
rd<-na.omit(100*diff(log(data)))#eliminate miss value
summary(rd)
plot(rd)
# monthly return is the sum of daily returns in a month
# rm=return monthly=100*diff_month(log(data))=100*ln(data@t)-ln(data@t-m)=100*ln(data@t/data@t-m)=sum r
# diff can be eliminate
# 月收益率rm
rm<-ts(apply.monthly(rd,sum),start=c(1997,7),frequency=12)
plot(rm)
monthplot(rm)#各月的年同比数据
seasonplot(rm,s=12,year.labels=T,col=rainbow(12))#各年的月度数据曲线
lag.plot(rm,12)#滞后相关图
Acf(rm,60,type="correlation")#ACF,lag60,if it cut at p order then it is a MA(p)
Acf(rm,60,type="partial")#PACF,lag60,if it cut at p order then it is a AR(p)
Acf(rm,200,type="correlation")#may be MA(107)
Acf(rm,150,type="partial")#may be AR(77)
```

#ACF & PACF
1. Autocorrelation function     
$$ACF=\rho_k = \frac{\gamma_k}{\gamma_0}=\frac{cov(X_t,X_{t-k})}{\sigma^{2}(X_t)}$$      
It is a decreasing function of k                
The speed of decreasing of stationary series is faster then nonstationary series        
2. Partial autocorrelation function     
$$
PACF=\varphi_{kk}\\
\textrm{ solve regression } \\
X_t=\varphi_{k1} X_{t-1}+\varphi_{k2}X_{t-2}+\varphi_{k3}X_{t-3}+\dots +\varphi_{kk}X_{t-k}+\epsilon_t
$$

|  模型   |           ACF           |               PACF                |   
|---------|-------------------------|-----------------------------------|       
|白噪声   |$\rho_k=0$               |$\rho_k=0$                         |       
|AR(p)    |衰减趋于0(几何形或震荡型)|p阶后截尾:$\rho_k^{*}=0,k>q$       |             
|MA(q)    |q阶后截尾:$\rho_k=0,k>q$ |衰减趋于0(几何型或震荡型)          |         
|ARMA(p,q)|q阶后衰减趋于0(几何型或震荡型)|p阶后衰减趋于0(几何型或震荡型)|         

# monthly volatility is the sum of daily squared returns in a month
月度风险是当月日回报率平方和
```{r}
rv<-ts(apply.monthly(rd^2,sum),start=c(1997,7),frequency=12)#transform return daily into sum of rd^2 monthly
plot(rv)
#每月的各年同比
monthplot(rv)
#每年的各月趋势
seasonplot(rv,s=12,year.labels=T,col=rainbow(12))
lag.plot(rv,12)
Acf(rv,60,type="correlation")
Acf(rv,60,type="partial")
Acf(rv,200,type="correlation")#may be MA(97)
Acf(rv,60,type="partial")#may be AR(5)
```
