---
title: "The Economist's Big Mac Price Index 1-8"
author: "Elara"
date: "2016年5月8日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Economist's Big Mac Price 
Historical Data from 1986-9-6 to 2016-1-7
Cross Sections and Time Series with Gaps
Unbalanced Panel Data
```{r}
# install.packages("plm")
library(plm)

setwd("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case1\\")
# load previously saved sorted (date) dataset
load("bmdatas.RData")   
class(bmdatas)
summary(bmdatas)

price<-sapply(split(bmdatas$dollar_price,bmdatas$id),mean)
ex<-sapply(split(bmdatas$dollar_ex,bmdatas$id),mean)

# Set data as panel data
bmdata <- plm.data(bmdatas, index=c("id","date"))
class(bmdata)
summary(bmdata)

# panel data analysis
Y<-bmdata$dollar_price
X<-bmdata$dollar_valuation
# X<-100*(bmdata$dollar_ppp/bmdata$dollar_ex - 1)  # dollar_valuation

# F test for fixed effects versus OLS
pFtest(Y ~ X, data=bmdata, effect="individual")
pFtest(Y ~ X, data=bmdata, effect="time")

# LM test for random effects versus OLS
plmtest(Y ~ X, data=bmdata, effect="individual", type="bp")
plmtest(Y ~ X, data=bmdata, effect="time", type="bp")

# Pooled OLS estimator
pooling <- plm(Y ~ X, data=bmdata, model= "pooling")
summary(pooling)

# Between estimator
between <- plm(Y ~ X, data=bmdata, model= "between")
summary(between)

# First differences estimator
firstdiff <- plm(Y ~ X, data=bmdata, model= "fd")
summary(firstdiff)

# Fixed effects or within estimator
fixed <- plm(Y ~ X, data=bmdata, model= "within")
summary(fixed)
# extract fixed effects
summary(fixef(fixed))

# Random effects estimator
random <- plm(Y ~ X, data=bmdata, model= "random") # use random.method="swar"
summary(random)

# Hausman test for fixed versus random effects model
phtest(random,fixed)

library("lmtest")
coeftest(fixed,vcovHC)
coeftest(random,vcovHC)

plot(X,Y)
abline(pooling)
abline(random,col="red")
# abline(fixed,col="blue") # need to adjust