---
title: "The Economist's Big Mac Price Index 1-7"
author: "Elara"
date: "2016年5月8日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Economist's Big Mac Price 
Historical Data from 1986-9-6 to 2015-7-21
Cross Sections and Time Series with Gaps
Unbalanced Panel Data
```{r}
# Using readxl package to read Excel spreadsheet
# install.packages("readxl")
setwd("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case1\\")
# load previously saved RData file
load("bmdatas.RData")
# test of equal variance across countries
bartlett.test(bmdatas$dollar_price,bmdatas$id)

# individual country's price time series data
bm_us<-subset(bmdatas,id=="United States")
bm_china<-subset(bmdatas,id=="China")
bm_sweden<-subset(bmdatas,id=="Sweden")

Y<-bm_china
# test for serial correlation
#求出1-5阶滞后的系列相关性
for (l in 1:5) {
  t<-Box.test(Y$dollar_price,lag=l)
  print(t)
}
#画acf图
acf(Y$dollar_price)
#画pacf图
pacf(Y$dollar_price)

# time series analysis,参数用order AP(P) 差分阶数 MA(q) 
arima(Y$dollar_price,order=c(1,1,0))

# similar analysis for other variable dollar_valuation



