---
title: "The Economist's Big Mac Price Index1.1"
author: "Elara"
date: "2016年5月7日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Economist's Big Mac Price Index1.1
The Economist's Big Mac Price (2016-1-7, 56 countries)   
Data Preparation  
Need data on local_price, dollar_ex (others can be computed)  
Using readxl package to read Excel spreadsheet  
```{r}
# install.packages("readxl")
setwd("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case1\\")
library(readxl)
args(read_excel)
```
read_excel can not read from url as "http://web.pdx.edu/~crkl/ec510/data/..."    
bm201601<-read_excel("http://infographics.economist.com/2016/databank/BMfile2000toJan2016.xls",sheet=1)  
try download.file first, then read it from there  
download.file("http://infographics.economist.com/2016/databank/BMfile2000toJan2016.xls","bm2016.xls",mode="wb")  
the following data file has been modified to include gdp_per_capita and Country=id  
read不能直接读网页，用download.file("地址","文件名")  
```{r}
download.file("http://web.pdx.edu/~crkl/WISE2016/data/BMfile2000-2016.xls","bm2016.xls",mode="wb")
bm201601<-read_excel("bm2016.xls",sheet=1)
names(bm201601)
dim(bm201601)
```
dollar_price=local_price/dollar_ex  
dollar_ppp=local_price/local_price["United States"]  
dollar valuation=100*(dollar_ppp-dollar_ex)/dollar_ex  
(-% undervaluation, +% overvaluation)  

# price in US$  
```{r}
bmp<-bm201601$dollar_price
hist(bmp)
hist(bmp,freq=F)
lines(density(bmp))
```  

# dollar valuation (-% undervaluation, +% overvaluation)  
```{r}
plot(bm201601$dollar_valuation,main="Valuation Against U.S.Dollar, January 2016",
     xlab="Country Index",ylab="U.S.$ Valuation in %")
abline(h=0)
text(bm201601$dollar_valuation,labels=bm201601$id,cex=0.5)
```

# gdp-adjusted valuation  
```{r}
Y<-bm201601$dollar_valuation-bm201601$dollar_adj_valuation
plot(Y,main="GDP-Adjusted Valuation, January 2016",
     xlab="Country Index",ylab="Valuation in %")
abline(h=0)
text(Y,labels=bm201601$id,cex=0.5)
```


