---
title: "Wine Price in Vancouver BC, Canada 2-2"
author: "Wang ZeXian"
date: "2016年5月10日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

TOTAL WEEKLY SALES OF IMPORT AND DOMESTIC NON VQA RED & WHITE TABLE WINE 
WITHIN MUNICIPALITY OF VANCOUVER IN UNITS AND LITRES
FROM WEEK ENDING APRIL 4, 2009 TO WEEK ENDING MAY 28, 2011
Using readxl package to read Excel spreadsheet: read_excel
Using qplot() of ggplot2 package for data visualization
#导入数据
```{r}
# install.packages("ggplot2")
# install.packages("readxl")
# library(readxl) 
# wine<-read_excel("C:/Course16/EC510/data/Vancouver_Non_VQA_Sls_Apr1toMay31_rev.xlsx",na="NA",skip=5)
# wine<-read_excel("C:/Course16/EC510/data/Vancouver_Non_VQA_Sls_Apr1toMay31.xlsx",sheet=2)
library(readr) 
# wine<-read_csv("http://web.pdx.edu/~crkl/WISE2016/data/Vancouver_Non_VQA_Sls_Apr1toMay31.csv",na="NA")
load("C:\\Users\\44180\\Documents\\R-in-SOE\\R\\big data\\case2\\wine.RData")
dim(wine)
names(wine)
summary(wine)
```

#数据整理
```{r}
#产地
where<-wine$`Store Category Sub Name` 
#substr(where,1,6).提取每一个字符串的前6个字母
head(where)
head(substr(where,1,6))
#数据中同一个国家有不同产地
where[226]
where[876]
#合并法国的产地都为FRANCE
where<-ifelse(substr(where,1,6)=="FRANCE","FRANCE",where)  
#合并德国的产地都为GERMANY
where<-ifelse(substr(where,1,7)=="GERMANY","GERMANY",where) 
big5<-c("AUSTRALIA","UNITED STATES","FRANCE","ITALY","OTHERS")
#is.element(where,big5) 返回where中元素是否是big5中的元素
head(is.element(where,big5))
#不是big5的就用OTHERS替代
where5<-ifelse(!is.element(where,big5),"OTHERS",where)
head(where5,100)
#加入where5作为新的因子
wine$where<-factor(where5)
#酒名
wine$what<-factor(wine$`Store Category Minor Name`)   # Red or White
#销量
wine$quantity<-wine$`Total Weekly Selling Unit`
```

#价格分析

```{r}
# price analysis
wine$price<-wine$`Current Display Price`
wine$price_level<-cut(wine$price,breaks=c(-Inf,50,100,1000,Inf),labels=c("Low","Medium","High","Expensive"))
#与Wine Price in Vancouver BC, Canada 2-1对比,此处简化
t1<-with(wine,table(where,price_level))
t1
barplot(t1)

t2<-with(wine,table(what,price_level))
t2
barplot(t2)

library(ggplot2)
#每种价格水平的具体价格箱线图
qplot(x=price_level,y=price,geom="boxplot",data=wine)
#分组散点图
qplot(x=price_level,y=price,geom="jitter",data=wine)
qplot(x=price_level,y=price,geom="point",data=wine)
#价格直方图
qplot(price,geom="histogram",data=wine)
summary(wine$price)
```

#缩小数据

```{r}
# reduce dataset to consider only the case of price<=100
varsel<-c("what","where","quantity","price","price_level")  
wine100<-subset(wine,price<100 & quantity>0,varsel) 
rm(wine)
```

#图像

```{r}
qplot(x=where,y=price,geom="boxplot",data=wine100)
qplot(x=what,y=price,geom="jitter",data=wine100)
#频数直方图
qplot(price,data=wine100) 
#频数折线图
qplot(price,geom="freqpoly",data=wine100)
#密度图(频率折线图)
qplot(price,geom="density",color=I("blue"),data=wine100)
#分组填色直方图$密度图
qplot(price,geom="histogram",fill=what,data=wine100)
qplot(price,geom="histogram",fill=where,data=wine100)
qplot(price,geom="density",fill=where,data=wine100)

# Using Facets谱图
qplot(price,geom="bar",fill=what,data=wine100)
qplot(price,geom="bar",data=wine100) + facet_grid(what ~ .)
qplot(price,geom="bar",data=wine100) + facet_grid(. ~ what)
qplot(price,geom="bar",data=wine100) + facet_grid(where ~ .)
qplot(price,geom="bar",data=wine100) + facet_grid(. ~ where)
```

#需求分析

```{r}
# demand analysis
model1<-lm(quantity~price,data=wine100)
summary(model1)
#散点图,透明度1/3
qplot(price,quantity,data=wine100,alpha=I(1/3))
qplot(price,quantity,data=wine100,alpha=I(1/3),col=what)
qplot(price,quantity,data=wine100,alpha=I(1/3),col=where)
#qplot(price,quantity,data=wine100,alpha=I(1/3),col=where,geom=c("point","smooth"),method="lm")
qplot(price,quantity,data=wine100,alpha=I(1/3),col=where,geom=c("point"))+stat_smooth(method="lm")
qplot(price,quantity,data=wine100,alpha=I(1/3),col=where,geom=c("point"))+stat_smooth(method="lm",color="darkblue")+facet_grid(where ~ .)
# log-model could perform better
model2<-lm(log(quantity)~log(price),data=wine100)
summary(model2)
#log="xy",x和y都做对数变化
qplot(price,quantity,data=wine100,alpha=I(1/3),log="xy")
qplot(price,quantity,data=wine100,alpha=I(1/3),col=what,log="xy")
qplot(price,quantity,data=wine100,alpha=I(1/3),col=where,log="xy")
#qplot(price,quantity,data=wine100,alpha=I(1/3),geom=c("point","smooth"),method="lm",log="xy")
qplot(price,quantity,data=wine100,alpha=I(1/3),geom=c("point"),log="xy")+stat_smooth(method="lm")
#qplot(price,quantity,data=wine100,alpha=I(1/3),col=what,geom=c("point","smooth"),method="lm",log="xy")
qplot(price,quantity,data=wine100,alpha=I(1/3),col=what,geom=c("point"),log="xy")+stat_smooth(method="lm")
#qplot(price,quantity,data=wine100,alpha=I(1/3),geom=c("point","smooth"),method="lm",log="xy") + facet_grid(what ~ .)
qplot(price,quantity,data=wine100,alpha=I(1/3),geom=c("point"),log="xy") + facet_grid(what ~ .)+stat_smooth(method="lm")
#qplot(price,quantity,data=wine100,alpha=I(1/3),geom=c("point","smooth"),method="lm",log="xy") + facet_grid(where ~ .)
qplot(price,quantity,data=wine100,alpha=I(1/3),geom=c("point"),log="xy") + facet_grid(where ~ .)+stat_smooth(method="lm")


# xlim, ylim, xlab, ylab, main may be used as in basic plot